Predicting Covid Data

### 2022-12-05 chien_sim.Rmd
update

```{r cars}
covid <- read_csv("owid-covid-data.csv")
```

```{r}
china <- covid %>% 
  filter(location=="China") %>% 
  filter(date <= "2020-03-28") %>% 
  filter(date >= "2020-01-22") %>% 
  select(date, total_cases, total_deaths) %>% 
  mutate(days = c(1:67))
```

@@

```{r}
italy <- covid %>% 
  filter(location=="Italy") %>% 
  filter(date <= "2020-04-21") %>%
  filter(date >= "2020-02-15") %>%
  select(date, total_cases, total_deaths) %>% 
  mutate(days = c(1:67))
italy[45:67,] <- NA
```

```{r}
ggplot() +
  geom_point(aes(x = date, y = total_cases), data = italy)
#+ geom_smooth(aes(x = date, y = total_cases), data=italy)
ggplot() +
  geom_point(aes(x = date, y = total_deaths), data = italy)
#+ geom_smooth(aes(x = date, y = total_deaths), data=italy)
```

```{r}
x <- 1:67
ggplot(italy,aes(x)) +
  geom_point(aes(x = days, y = total_cases)) +
  stat_function(fun=function(x) 100000 + 100000 * erf(0.07*x - 3))

ggplot(italy,aes(x)) +
  geom_point(aes(x = days, y = total_deaths)) +
  stat_function(fun=function(x) 11000 + 11000 * erf(0.068*x - 3))
```

>>>>

```{r}
data <- read_csv("owid-covid-data.csv")
```

```{r}
china <- data %>% 
  filter(location=="China") %>% 
  filter(date <= "2020-03-28") %>% 
  filter(date >= "2020-01-22") %>% 
  select(date, total_cases, total_deaths) %>% 
  mutate(days = c(1:67))
```

@@

## ITALY

```{r}
italy <- data %>% 
  filter(location=="Italy") %>% 
  filter(date <= "2020-04-21") %>%
  filter(date >= "2020-02-15") %>%
  select(date, total_cases, total_deaths) 
italy[45:67,] <- NA
italy <- italy %>% mutate(days = c(1:67))
```

```{r}
ggplot(italy, aes(x = days, y = total_cases)) + 
    geom_point() + 
    labs(title = "Total Cases in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Cases")
ggplot(italy, aes(x = days, y = total_deaths)) +
    geom_point() + 
    labs(title = "Total Deaths in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Deaths")

```

```{r}
# # without nls fxn
# ggplot(italy,aes(x)) +
#   geom_point(aes(x = days, y = total_cases)) +
#   stat_function(fun=function(x) 100000 + 100000 * erf(0.07*x - 3))
# 
# ggplot(italy,aes(x)) +
#   geom_point(aes(x = days, y = total_deaths)) +
#   stat_function(fun=function(x) 11000 + 11000 * erf(0.068*x - 3))

```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
library(minpack.lm)
x <- 1:67

# with nls fxn
gauss_fn <- function(x,a,b,c,d) {a+b*erf(c*x-d)}

c <- summary(nlsLM(total_cases~gauss_fn(days,a,b,c,d), data=italy, start=list(a=50000,b=50000,c=.01,d=1)))$parameters
d <- summary(nlsLM(total_deaths~gauss_fn(days,a,b,c,d), data=italy, start=list(a=10000,b=10000,c=.01,d=1)))$parameters

cf_italy <- function(x) c[1]+c[2]*erf(c[3]*x-c[4])
df_italy <- function(x) d[1]+d[2]*erf(d[3]*x-d[4])

ggplot(italy, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cf_italy, color="#d33636", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Cases in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Cases")

ggplot(italy, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=df_italy, color="#1b95ff", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Deaths in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Deaths")
```

## montecarlo sim make matrix

```{r}
n <- 44
m <- 150
sigma <- 0.1
mu <- 1
gau_sample <- matrix(rnorm(n*m, mean=mu, sd=sigma), nrow = m, ncol = n)

italy_mod <- italy[1:44,]
italy_mod[c(1:6),"total_deaths"] = 0

model_ca <- c()
for (i in 1:n) {
    model_ca <- c(model_ca, italy_mod$total_cases[i]*gau_sample[,i])
}
model_ca <- matrix(model_ca, nrow = m, ncol = n)

model_fa <- c()
for (i in 1:n) {
    model_fa <- c(model_fa, italy_mod$total_deaths[i]*gau_sample[,i])
}
model_fa <- matrix(model_fa, nrow = m, ncol = n)
```

## Day of flex (trial)

```{r}
# trial 1
test <- model_fa[11,]
model_try <- nlsLM(test ~ gauss_fn(days,a,b,c,d), 
                 data=italy_mod, 
                 start=list(a=50000,b=50000,c=.01,d=1))
est_try <- summary(model_try)$parameters
est_try <- est_try[, 1]
names(est_try) <- NULL
# fn_try <- function(x, ests) {ests[1]  + ests[2] * erf(ests[3] * x - ests[4])}

erf_deriv <- function(x) {2/sqrt(pi)*exp(-x^2)}

fn_try_deriv <- bquote(.(est_try[2]) * 2/sqrt(pi) * exp(-(.(est_try[3]) * x - .(est_try[4]))^2))
fn_try_deriv <- D(fn_try_deriv,'x')
x=seq(1,67,length.out=6700)
tmp <- eval(fn_try_deriv)
min(x[tmp < 0])
```

##

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# cases
flex_ca <- rep(NA, m)
for (i in 1:m) {
  per_m <- model_ca[i, ]
  model_try <-nlsLM(per_m ~ gauss_fn(days, a, b, c, d),
                  data = italy_mod,
                  start = list(a=50000,b=50000,c=.01,d=1))
  get_coef <- summary(model_try)$parameters
  get_coef <- get_coef[, 1]
  names(get_coef) <- NULL
  # Derivative of Gauss Error Function: 2 / sqrt(pi) * exp(-x ^ 2)
  fn_try_deriv <-
    bquote(.(get_coef[2]) * 2 / sqrt(pi) * exp(-(.(get_coef[3]) * x - .(get_coef[4])) ^ 2))
  fn_try_deriv <- D(fn_try_deriv, 'x')
  x = seq(1, 67, length.out = 6700)
  tmp <- eval(fn_try_deriv)
  flex_ca[i] <- min(x[tmp < 0])
}
mean(flex_ca)
sd(flex_ca)
```

```{r}
# fatalities
flex_fa <- rep(NA, m)
for (i in 1:m) {
  per_m <- model_fa[i, ]
  model_try <-nlsLM(per_m ~ gauss_fn(days, a, b, c, d),
                  data = italy_mod,
                  start = list(a=15000,b=15000,c=.01,d=1))
  get_coef <- summary(model_try)$parameters
  get_coef <- get_coef[, 1]
  names(get_coef) <- NULL
  # Derivative of Gauss Error Function: 2 / sqrt(pi) * exp(-x ^ 2)
  fn_try_deriv <-
    bquote(.(get_coef[2]) * 2 / sqrt(pi) * exp(-(.(get_coef[3]) * x - .(get_coef[4])) ^ 2))
  fn_try_deriv <- D(fn_try_deriv, 'x')
  x = seq(1, 67, length.out = 6700)
  tmp <- eval(fn_try_deriv)
  flex_fa[i] <- min(x[tmp < 0])
}
flex_fa <- flex_fa[-c(12,112,124,133)]
mean(flex_fa)
sd(flex_fa)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
#output
print(mean(flex_fa))
print(sd(flex_fa))
print(mean(flex_ca))
print(sd(flex_ca))
```

```{r}
#plot
ggplot(italy, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cf_italy, color="#d33636", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Cases in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Cases") + 
  geom_vline(xintercept = mean(flex_ca), color="#d33636", size=1.5, linetype="dashed")

ggplot(italy, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=df_italy, color="#1b95ff", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Deaths in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Deaths") + 
  geom_vline(xintercept = mean(flex_fa), color="#1b95ff", size=1.5, linetype="dashed")
```

### 2022-12-05 main.rmd
update

* In this study by Ignazio Ciufolini and Antonio Paolozzi, they use a Gauss Error function that approximates the distribution of the cumulative positive cases of COVID-19 in China
  * They have then use the function to approximate the number of cumulative positive cases in Italy
  * Then performed a 150 Monte Carlo simulations to predict the peak and valley of the number of daily positive cases and fatalities
* This study do not take into account factors such as daily nasopharyngeal swabs, medical, social distancing, etc.

## Cont.

* The number of cumulative diagnosed positive cases and fatalities in China approximates a gauss error function: $a+b*erf(c*x-d)$.
* Predicted the potential number of cases in Italy.
* Predicted the day in which the peak of the number of daily cases in Italy occurs.

* create a random matrix : m X n
  * m = number of random outcomes (*chosen to be 150*)
  * n = number of observed days (*n = 1,...,j*)
* each # in the matrix is a Gaussian distribution with mean = 1 and sigma = 0.1
  * multiply each column (*j*) by the # of total cases (or deaths) that corresponds to each day *j* **#FIXME#**
- add more...

@@

## Plot of Diagnosed Positive Cases

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# extract Italy data
italy <- data %>% 
  filter(location=="Italy") %>% 
  filter(date <= "2020-04-21") %>%
  filter(date >= "2020-02-15") %>%
  select(date, total_cases, total_deaths) %>% 
  mutate(days = c(1:67))
italy[45:67,] <- NA

ggplot() +
  geom_point(aes(x = date, y = total_cases), data = italy)
#+ geom_smooth(aes(x = date, y = total_cases), data=italy)
```

## Plot of Fatalities

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  geom_point(aes(x = date, y = total_deaths), data = italy)
#+ geom_smooth(aes(x = date, y = total_deaths), data=italy)
```

## Plot of Diagnosed Positive Cases with Gauss Error Function

```{r, echo=FALSE, message=FALSE, warning=FALSE}
x <- 1:67
ggplot(italy,aes(x)) +
  geom_point(aes(x = days, y = total_cases)) +
  stat_function(fun=function(x) 100000 + 100000 * erf(0.07*x - 3))
```

## Plot of Fatalities with Gauss Error Function

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(italy,aes(x)) +
  geom_point(aes(x = days, y = total_deaths)) +
  stat_function(fun=function(x) 11000 + 11000 * erf(0.068*x - 3))
```

>>>>

* The number of cumulative diagnosed positive cases and fatalities in China approximates a gauss error function:  
$a+b*erf(c*x-d)$.
* Used the function to approximate the number of cumulative positive cases in Italy
* Performed a 150 Monte Carlo simulations to predict the day in which the peak of the number of daily cases in Italy occurs.
* This study do not take into account factors such as daily nasopharyngeal swabs, medical, social distancing, etc.

@@

## Initial Data for Cases in Italy

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# extract Italy data
italy <- data %>% 
  filter(location=="Italy") %>% 
  filter(date <= "2020-04-21") %>%
  filter(date >= "2020-02-15") %>%
  select(date, total_cases, total_deaths)
  #ref: %>% mutate(days = c(1:67))
italy[45:67,] <- NA
italy <- italy %>% mutate(days = c(1:67))

#plot
ggplot(italy, aes(x = days, y = total_cases)) + 
    geom_point() + 
    labs(title = "Total Cases in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Cases")
```

## Initial Data for Fatalities in Italy

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#plot
ggplot(italy, aes(x = days, y = total_deaths)) +
    geom_point() + 
    labs(title = "Total Deaths in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Deaths")
```

## Fitting Gauss Error Function to Positive Cases in Italy

```{r, echo=FALSE, message=FALSE, warning=FALSE}
x <- 1:67

c <- summary(nlsLM(total_cases~gauss_fn(days,a,b,c,d), data=italy, start=list(a=50000,b=50000,c=.01,d=1)))$parameters
d <- summary(nlsLM(total_deaths~gauss_fn(days,a,b,c,d), data=italy, start=list(a=10000,b=10000,c=.01,d=1)))$parameters

cf_italy <- function(x) c[1]+c[2]*erf(c[3]*x-c[4])
df_italy <- function(x) d[1]+d[2]*erf(d[3]*x-d[4])

#plot
ggplot(italy, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cf_italy, color="#d33636", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Cases in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Cases")
```

## Fitting Gauss Error Function to Fatalities in Italy

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#plot
ggplot(italy, aes(x = days, y = total_deaths)) +
    geom_point() + 
    stat_function(fun=df_italy, color="#1b95ff", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Deaths in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Deaths")
```

* create a random matrix : m X n
  * m = number of random outcomes (*chosen to be 150*)
  * n = number of observed days (*n = 1,...,j*)
* each # in the matrix is a Gaussian distribution with mean = 1 and sigma = 0.1
  * multiply each column (*j*) by the # of total cases (or deaths) that corresponds to each day *j*
- add more...

- The mean of the day of flex for cases is 40.59 days and the standard deviation is 5.222832 days.
- The mean of the day of flex for fatalities is 43.46 days and the standard deviation is 5.228381 days.

## 150 Monte Carlo simulations plotted

"Each red dot corresponds to the day of occurrence of the flex obtained with each of the 150 Monte Carlo simulations"

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
n <- 44
m <- 150
sigma <- 0.1
mu <- 1
gau_sample <- matrix(rnorm(n*m, mean=mu, sd=sigma), nrow = m, ncol = n)

italy_mod <- italy[1:44,]

model_ca <- c()
for (i in 1:n) {
    model_ca <- c(model_ca, italy_mod$total_cases[i]*gau_sample[,i])
}
model_ca <- matrix(model_ca, nrow = m, ncol = n)

model_fa <- c()
for (i in 1:n) {
    model_fa <- c(model_fa, italy_mod$total_deaths[i]*gau_sample[,i])
}
model_fa <- matrix(model_fa, nrow = m, ncol = n)
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# cases
flex_ca <- rep(NA, m)
for (i in 1:m) {
  per_m <- model_ca[i, ]
  model_try <-nlsLM(per_m ~ gauss_fn(days, a, b, c, d),
                  data = italy_mod,
                  start = list(a=50000,b=50000,c=.01,d=1))
  get_coef <- summary(model_try)$parameters
  get_coef <- get_coef[, 1]
  names(get_coef) <- NULL
  # Derivative of Gauss Error Function: 2 / sqrt(pi) * exp(-x ^ 2)
  fn_try_deriv <-
    bquote(.(get_coef[2]) * 2 / sqrt(pi) * exp(-(.(get_coef[3]) * x - .(get_coef[4])) ^ 2))
  fn_try_deriv <- D(fn_try_deriv, 'x')
  x = seq(1, 67, length.out = 6700)
  tmp <- eval(fn_try_deriv)
  flex_ca[i] <- min(x[tmp < 0])
}
mean(flex_ca)
sd(flex_ca)
```

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  geom_point(aes(x=1:m, y=flex_ca), color="red") +
  geom_hline(yintercept=mean(flex_ca), color="blue") +
  labs(x = "Number of Monte Carlo simulation", y = "Day of flex, from February 15")
```

## Applying the day of flex indicator to the plot (Cases)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#plot(cs)
ggplot(italy, aes(x = days, y = total_cases)) + 
    geom_point() + 
    stat_function(fun=cf_italy, color="#d33636", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Cases in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Cases") + 
  geom_vline(xintercept = mean(flex_ca), color="#d33636", size=1.5, linetype="dashed")
```

## Applying the day of flex indicator to the plot (Fatalities)

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# fatalities
flex_fa <- rep(NA, m)
for (i in 1:m) {
  per_m <- model_fa[i, ]
  model_try <-nlsLM(per_m ~ gauss_fn(days, a, b, c, d),
                  data = italy_mod,
                  start = list(a=15000,b=15000,c=.01,d=1))
  get_coef <- summary(model_try)$parameters
  get_coef <- get_coef[, 1]
  names(get_coef) <- NULL
  # Derivative of Gauss Error Function: 2 / sqrt(pi) * exp(-x ^ 2)
  fn_try_deriv <-
    bquote(.(get_coef[2]) * 2 / sqrt(pi) * exp(-(.(get_coef[3]) * x - .(get_coef[4])) ^ 2))
  fn_try_deriv <- D(fn_try_deriv, 'x')
  x = seq(1, 67, length.out = 6700)
  tmp <- eval(fn_try_deriv)
  flex_fa[i] <- min(x[tmp < 0])
}
flex_fa <- flex_fa[-c(12,112,124,133)]
mean(flex_fa)
sd(flex_fa)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#plot(ft)
ggplot(italy, aes(x = days, y = total_deaths)) +
    geom_point() +
    stat_function(fun=df_italy, color="#1b95ff", size=2, linetype="solid") +
    scale_x_continuous(breaks = seq(0, 67, 10)) +
    labs(title = "Total Deaths in Italy from 2/15/2020 to 3/29/2020 (44 days)", x = "Days", y = "Total Deaths") +
  geom_vline(xintercept = mean(flex_fa), color="#1b95ff", size=1.5, linetype="dashed")
```
